Summarize distribution of LR statistic under null

### 2016-06-08 analysis/referee.response.Rmd

---
title: "Response to referees"
author: "Matthew Stephens"
date: 2016-03-21
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

```{r chunk-options, include=FALSE}
source("chunk-options.R")
```

# Referee 1

I agree with a lot of what the referee says, although there are some points of disagreement. The most fundamental disagreement regards Figure 1. This Figure shows
not the detectable effects (as the referee believed) but the decomposition
of *all* effects into null and alternative components. Since this seems
to be an important misunderstanding I have revised the text to try to make it clearer how this figure was derived. Specifically the revised text says:

“We used each of the methods qvalue, locfdr, mixfdr and ash to decompose the $z$ scores ($z_j = \hat{\beta}_j$), or their corresponding p values, into null and alternative components. Here we are using the fact that these methods all provide an estimate of the lfdr for each observation, which implies such a decomposition; specifically the average lfdr within each histogram bin estimates the fraction of observations in that bin that come from the null vs the alternative component.”

Since the referee objected to the presentation of the ZA and UA as opposing assumptions I have removed all explicit reference to the ZA. (This also 
solves the issue that I had defined the ZA differently than Efron’s original definition — see below — which may have created confusion.) 
Instead, as the referee 
suggested, I have presented the UA as an additional assumption “not made by other methods”, and - as the editor suggested —
I have focussed on contrasting the behavior of ash with the behavior 
of the other methods (which, as Figure 1 demonstrates, is to create a hole at 0 in the distribution of the alternative Z scores) rather than contrasting their assumptions.

However, it seems unsatisfactory to merely point out this behavioral
difference between methods without giving any explanation for the reason behind it.
So I have provided my best attempt to give the reason. Essentially the reason is that the existing methods (locfdr and qvalue), when estimating $\pi_0$, make the assumption that  *all* p values near 1 are null, or *all* z scores near 0 are null. 
(This was how I defined the ZA in the original paper; however this is stronger than Efron’s original statement of the ZA, which the referee - understandably - uses in his/her report, that ``most” z scores near 0 are null. So I don’t refer to this
as the ZA any more.) This assumption necessarily creates a hole in the
implied distribution of z scores under the alternative, and explains the behavior.

To underscore that this behavior is general, and not just a feature of the particular
simulation I did in Figure 1, I provide results using qvalue under the setting
suggested by the referee at (http://stephenslab.github.io/ash/analysis/referee_uaza.html).

## Minor Concerns

- I added the citation, thank you.

- I took up the suggestion to define FSR-hat and the s-value, thank you.

- I agree that returning the length J results as a dataframe would have been better, but
unfortunately changing this would break other packages that depend on ashr. So instead I have provided a method `as.data.frame` to extract the most important components of the ash object into a dataframe.

- I fixed the typo, thank you.

# Referee 2

## Major Concern

The referee says: "it would be interesting to
include distributions that favor other models, such as a bimodal distribution with no mass at zero, an assymmetrical distribution,. . . The unimodality is a strong assumption so it would be good to see how anti-conservative the results would get under different alternative distributions."

I found this puzzling since both an assymetrical distribution and a strongly bimodal distribution are already included in the simulation results (Figure 2, Figure 3, Table 1). I believe these existing results address the referee comment. I also want to emphasise that the unimodal assumption is actually less strong than the assumptions usually made in variable selection for regression, which is directly analogous to the fdr context. For example, it is common to assume that effects are a mixture of a point mass at zero and a single normal, which is considerably more restrictive than the general unimodal assumption.

However, to help reassure the referee, I have added another example to illustrate an assymetric situation where the unimodal-at-zero assumption is badly contradicted [here](http://stephenslab.github.io/ash/analysis/efron.fcr.html). As
the referee will see, the resulting estimates of the lfdr from ashr are
reasonably close to the true values (and also to those from locfdr).

## Minor Concern

- Is there a way to model the distribution as a mixture of both uniforms and normals? If not: why?

This is straightforward in principle. However, it would require 
non-trivial code modifications because of the way I implemented the mixture class (assuming that all components  come from the same family.) 
Another way to achieve an assymetric distribution without
the "hard" tails of the uniform might be to use a mixture of truncated normals. I have a student investigating this option. However, the uniform mixture has the advantage of allowing not only the assymetry, but also the $t$ likelihood. Although the paper
does not focus on this, I believe this is likely to be an important issue in practice.
So I suspect that the mixture of uniforms is often going to be the method of choice in practice.

- How would the model perform with moderate correlations?

There seems to be no reason to think that
moderate correlations among tests will have a substantial detrimental effect. 
But I do not want
to focus on this because I think it will give the reader the wrong message:
I think that users of both ashr and other fdr methods
should very much worry about correlations in practice (which I believe will
often be large, and not moderate).

- In equation 3, the null component (I think) is represented as delta0, but delta0 hasn’t been defined. What is assumed for the distribution of delta0?

Thank you for noticing this. I now define $\delta_0$ (which represents a point mass on 0).

- Figure 3: the truth is hard to distinguish from the different methods. And in general this figure is hard to read.

I agree. I have completely rethought this figure, which now has fewer, larger,
panels, and compares just two  methods (ash.hu and the NPMLE) against the truth on a few examples. The less important point - the effect of the penalty term on pi0 -is now illustrated in an additional figure included in the appendix.

- I particularly like the idea that the ordering of the tests differs from the classical p-values. Which influence does this have on the balance between sensitivity and specificity (for example with ROC curves)? It should change, but does it get better?

Thanks for the question. This is now addressed in the new Figure 4d.

## Typos and grammar

- page 14, line 20. The first mentioned pi0 and lfdr should have a hat.

Thanks - fixed.

## Session information

```{r info}
sessionInfo()
```

### 2016-06-08 analysis/index.Rmd
re-run analysis, finalize edits

*  [Illustrate Zero Assumption](makefig_FDReg.html)
*  [Illustatrate dilution effect of low-precision observations](make_GOODPOOR_figs.html)
*  [Adaptive Shrinkage](adaptive_shrinkage.html)
*  [Comparing models: EE vs ET models](CompareEEvsETmodel.html)
*  [Checking how qvalue behaves in simulation suggested by referee](referee_uaza.html)
*  [Example of Credible Interval behaviour when UA does not hold](efron.fcr.html)

>>>>

*  [Figures illustrating FDR](makefig_FDReg.html)
*  [Illustatrate dilution effect of low-precision observations](make_GOODPOOR_figs.html)
*  [Adaptive Shrinkage](adaptive_shrinkage.html)
*  [Comparing models: EE vs ET models](CompareEEvsETmodel.html)
*  [Checking how qvalue behaves in simulation suggested by referee](referee_uaza.html)
*  [Comparison of methods, and example of Credible Interval behaviour when UA does not hold](efron.fcr.html)
*  [Using ash to fit the NPMLE; comparison with REBayes](ash_npmle.html)

### 2016-06-08 analysis/summarize_dsc_znull.Rmd
re-run analysis, finalize edits

```{r}
ggplot(res %>% filter(logLR>0) %>% filter(grepl("ash.hu",method)) , aes(sample = logLR)) + facet_grid(scenario~method) +
  stat_qq(dist = qchisq, dparams=list(df=1)) + geom_abline() + ggtitle("qqplot under null vs chisq-1; line slope=1")
ggplot(res %>% filter(logLR>0) %>% filter(grepl("ash.u",method)) , aes(sample = logLR)) + facet_grid(scenario~method) +
  stat_qq(dist = qchisq, dparams=list(df=1)) + geom_abline(intercept=0,slope=0.5) + ggtitle("qqplot under null vs chisq-1; line slope=0.5")
ggplot(res %>% filter(logLR>0) %>% filter(grepl("ash.n",method)), aes(sample = logLR)) + facet_grid(scenario~method) +
  stat_qq(dist = qchisq, dparams=list(df=1)) + geom_abline(intercept=0,slope=0.5) + ggtitle("qqplot under null vs chisq-1; line slope=0.5")

>>>>

```{r}
ggplot(res %>% filter(logLR>0) %>% filter(grepl("ash.hu",method)) , aes(sample = logLR)) + facet_grid(scenario~method) +
  stat_qq(distribution = qchisq, dparams=list(df=1)) + geom_abline() + ggtitle("qqplot under null vs chisq-1; line slope=1")
ggplot(res %>% filter(logLR>0) %>% filter(grepl("ash.u",method)) , aes(sample = logLR)) + facet_grid(scenario~method) +
  stat_qq(distribution = qchisq, dparams=list(df=1)) + geom_abline(intercept=0,slope=0.5) + ggtitle("qqplot under null vs chisq-1; line slope=0.5")
ggplot(res %>% filter(logLR>0) %>% filter(grepl("ash.n",method)), aes(sample = logLR)) + facet_grid(scenario~method) +
  stat_qq(distribution = qchisq, dparams=list(df=1)) + geom_abline(intercept=0,slope=0.5) + ggtitle("qqplot under null vs chisq-1; line slope=0.5")
